Visualizing Algebraic Surfaces

### 2016-04-14 project/article/src/article.tex
pulled over the stuff on projective space, generalized to complex setting

@@

@@

\newcommand{\C}{\ensuremath{\mathbb{C}}}

@@

>>>>

\usepackage[backend=biber, style=alphabetic]{biblatex} 
\addbibresource{../bib/mat2000.bib}

@@

\usepackage[]{enumerate}

@@

\theoremstyle{definition}
\newtheorem{defn}{Definition}

\newcommand{\C}{\ensuremath{\mathbb{C}}}
\newcommand{\proj}{\ensuremath{\mathbb{P}}}

@@

    The way we visualize these surfaces in four variables is by \emph{fixing}
    one of the variables. This is called \emph{homogenizing} the polynomial
    $h(x, y, z)$. We first want to introduce some definitions.

    Quite informally, projective space formalizes the notion of parallell lines
    intersecting at infinity. In order to get a better understanding of the
    general projective space $\proj^n$ we will first construct the
    \emph{projective plane} $\proj^2(\C)$.

    \subsection{The Projective Plane}
    \label{sub:the_projective_plane}

    According to \cite{Wik16} there are three equivalent definitions:
    \begin{enumerate}
        \item The set of all lines in $\C^3$ passing through the origin $(0, 0,
            0)$. Every such line meets the \emph{sphere} of radius one centered
            in the origin exactly twice, say in $P = (x, y, z)$ and its
            antipodal point $(-x, -y, -z)$.

        \item\label{item:sphere} The points on the sphere $S^2$, where every point $P$ and its
            antipodal points are not distinguished. For example, the point $(1,
            0, 0)$ is identified with the point $(-1, 0, 0)$.

        \item The set of equivalence classes of $\C^3 \setminus \left\{ 0
            \right\}$ where two points $P$ and $P'$ are considered equivalent
            if and only if there is a non-zero $\lambda \in \C$ such that $P =
            \lambda P'$.
    \end{enumerate}

    The last definition is the one seemingly most used, and is therefore the
    one we will employ here. Note that the elements in the projective plane are
    equivalence classes of points in $\C^3$. We denote an element in
    $\proj^2(\C)$ as $\left[ x : y : z \right]$. These elements are commonly
    referred to as \emph{homogenous coordinates}. We state this formally in a
    definition.

    \begin{defn}[The projective plane]
        Let $\sim$ be the equivalence relation on $\C^3$ defined by
        \begin{equation}
            \notag
            (x, y, z) \sim (x', y', z') \iff (x', y', z') = \lambda(x, y, z),
        \end{equation}
        where $\lambda$ is some non-zero scalar in $\C$. We denote the
        equivalence class of $(x, y, z)$ as $\left[ x : y : z \right]$ and we
        define the \emph{projective plane} as the set
        \begin{equation}
            \notag
            \proj^2(\C) = \left\{ \left[ x : y : z \right] \mid (x, y, z) \in
            \C^3 \setminus \left\{ 0 \right\}\right\}.
        \end{equation}
    \end{defn}

    We can now look at an interesting subset of $\proj^2(\C)$, namely the set
    of points $\left[x : y : z \right]$ where $z \neq 0$. Remembering the
    identification from the above definition, these are equivalent to $\left[
    \frac{x}{z} : \frac{y}{z} : 1 \right]$ where $\lambda = 1 / z$. If we
    consider definition \cref{item:sphere} from above, this corresponds to the
    set of points on the northern hemisphere of $S^2$ not including the circle
    of intersection in the $(x, y)$-plane. Similarly, we can look at the set of
    points $\left[ x : y : 0 \right]$. This correponds, in $S^2$, to the circle
    of intersection in the $(x, y)$-plane.

    We can now take the notion of the
    projective plane and generalize this to $n$ dimensions.

    \subsection{Projective Space}
    \label{sub:projective_space}

    The general definition of projective $n$-space is completely anologous to
    the definition of the projective plane, the set of lines in $\C^{n+1}$
    passing through the origin, however we include a formal definition for
    completeness.

    \begin{defn}[The projective space of dimension $n$]
        Let $\sim$ be an equivalence relation on $\C^{n+1} \setminus \left\{ 0
        \right\}$ defined by
        \begin{equation}
            \notag
            \left( x_1, \ldots, x_{n+1} \right) \sim \left( x_1', \ldots,
            \x_{n+1}' \right) \iff \left( x_1', \ldots, x_{n+1}' \right) =
            \lambda \left( x_1, \ldots, x_{n+1} \right)
        \end{equation}
        where $\lambda$ is some non-zero complex number. We then denote the
        equivalence class of $\left( x_1, \ldots, x_{n+1} \right)$ by $\left[
        x_1 : \ldots : x_{n+1} \right]$ and define the \emph{projective space
        of dimension $n$} as the set
        \begin{equation}
            \notag
            \proj^n{\C} = \left\{ \left[ x_1 : \ldots : x_{n+1} \right] \mid
            \left( x_1, \ldots, x_{n+1} \right) \in \C^{n+1} \setminus \left\{
            0 \right\} \right\}.
        \end{equation}
    \end{defn}

\printbibliography
